Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-02-23 appendix.tex
I add subsection of co-regular of PQ1>. I also add explaination for clarity. Now Appendix C.2 is for this subsection.

For each such case, we construct a witness automata. Recall that in Section \ref{subsec:definition of co-regular}, we generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

@@

\subsection{Co-Regular of $R_{\textit{pr1}}$}
\label{subsec:co-regular of Rpr1}

Let us introduce the notion of gap for matched $\textit{put}$ and $\textit{rm}$ operations.

\begin{definition}[gap for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap for matched put and rm operations}

Let $h$ be a data-differentiated history and $o_1 = \textit{put}(x,\textit{pri}_x)$, $o_2 = \textit{rm}(x)$ be a pair of matched $\textit{put}$ and $\textit{rm}$ operation of $h$. We say that $h$ has a gap on operations $o_1$ and $o_2$, if there is a partition of the operations of $h$ into $L$, $M$ and $R$ satisfying:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $L \cup M$ has matched $\textit{put}$ and $\textit{rm}$ operations.

\item[-] no operation of $R$ happens before an operation of $L$, $o_1$, $M$ or $o_2$.

\item[-] no operation of $o_2$ happens before an operation of $L$, $o_1$ or $M$.

\item[-] no operation of $M$ happens before an operation of $L$ or $o_1$.

\item[-] no operation of $o_1$ happens before an operation of $L$.

\item[-] no operation in $L \cup M \cup R$ has priority $\textit{pri}_x$.
\end{itemize}
\end{definition}

Given a history $h$ and set $O$ of some operations in $h$. It is easy to see that there exists a sequence $l$ of operations of $O$, such that $l$ does not contradicts the happen before relation of $h$. Therefore, from now on, we can use this fact to directly generate such sequence from such $O$.

The following lemma reduce linearizable with respect to $\textit{MR}_{\textit{pr1}}$ into gap of matched $\textit{put}$ and $\textit{rm}$ operations.

\begin{restatable}{lemma}{GapEqualsLinforRpr1}
\label{lemma:Gap Equals Lin for Rpr1}
A data-differentiated history $h$ has a projection $h'$, such that $\textit{last}(h') = R_{\textit{pr1}}$ and $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$, if and only if there exists a projection $h''$ of $h$, such that $\textit{last}(h'') = R_{\textit{pr1}}$, $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is in $h''$, $\textit{pri}_x$ is the maximal priority of $h''$, and one the the following case holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] case $1$: $\textit{rm}(x)$ happens before $\textit{put}(x,\textit{pri}_x)$,
\item[-] case $2$: there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
\end{itemize}
\end{restatable}

@@

Let us prove the $\textit{only if}$ direction by contradiction. Assume that there exists such $h'$, while case $1$ and case $2$ do not hold for any projection of $h$. Since case $1$ does not hold for any projection of $h$, we know that $\forall y \in D$, $\textit{rm}(y)$ does not happens before $\textit{put}(y,\_)$ in $h$. Since case $2$ does not hold for any projection of $h$ and $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $h'$ has gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, while $\textit{pri}_x$ is the maximal priority in $h'$. Since $\textit{last}(h') = R_{\textit{pr1}}$, we can see that $\textit{pri}_x$ is larger than priority of all other $\textit{put}$ operations of $h$.

By the properties of the gap, there is partitions $L$, $M$ and $R$. Let $u$, $v$ and $w$ contain the operations of $L$, $M$ and $R$, respectively. It is easy to see that $h'$ is linearizable with respect to $l''=u \cdot \textit{put}(x,\textit{pri}) \cdot v \cdot \textit{rm}(x) \cdot w$ and $l'' \in \textit{MR}_{\textit{pr1}}$. This contradicts the fact that $h'$ does not linearizable to $\textit{MR}_{\textit{pr1}}$.

Let us begin to prove the $\textit{if}$ direction.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If case $1$ holds for $h''$, then $h \vert_{ \{ x \} }$ is the $h'$ as required.

\item[-] If case $2$ holds, assume by contradiction that, for every projection $h'$ of $h$, if $\textit{last}(h') = R_{\textit{pr1}}$, then $h' \sqsubseteq \textit{MR}_{\textit{pr1}}$. Since $h''$ is also a projection of $h$, we know that there exist sequences $u$, $v$ and $w$, such that $h'' \sqsubseteq u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and $\textit{pri}_x$ is larger than priority of $u \cdot v \cdot w$. Let $L$, $M$ and $R$ be the operations of $u$, $v$ and $w$, respectively. Then, the partitions $L$, $M$ and $R$ form a gap on operation $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. This contradicts the fact that there is no gap on $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$.
\end{itemize}

\begin{definition}[left-right constraint for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:left-right constraint for matched put and rm operations}
Given a data-differentiated history $h$, and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. The left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ is the graph $G$ where:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the items of $h$, to which we add a node,

\item[-] there is an edge from item $d_1$ to $x$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{put}(x,\textit{pri}_x)$ or $\textit{rm}(x)$ and $\textit{pri}_1 \neq \textit{pri}_x$,

\item[-] there is an edge from $x$ to item $d_1$, if $\textit{rm}(x)$ happens before $\textit{rm}(d_1)$ or $\textit{rm}(d_1)$ does not exists in $h$, and the priority of $d_1$ is not $\textit{pri}_x$,

\item[-] there is an edge from item $d_1$ to item $d_2$, if $\textit{put}(d_1,\textit{pri}_1)$ happens before $\textit{rm}(d_2,\textit{pri}_2)$, $\textit{pri}_1 \neq \textit{pri}_x$, and $\textit{pri}_2 \neq \textit{pri}_x$.
\end{itemize}
\end{definition}

Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. Let $\textit{LMSet}_1(h,x) = \{ o \vert$ $o$ is not an operation of $x$, the priority of $o$ does not equal the priority of $x$, and either it happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$, or there is an operation $o'$ with the same item of $o$, such that $o'$ happens before $\textit{put}(x)$ or $\textit{rm}(x)$ in $h$ $\}$. For each $i \geq 1$, let $\textit{LMSet}_{\textit{i+1}}(h,x) = \{ o \vert$ $o$ is not an operation of $x$ or items in $\textit{LMSet}_k(h,x)$ for each $k \leq i$, the priority of $o$ does not equal the priority of $x$, and either it happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$, or there is an operation $o''$ with the same item of $o$ and $o''$ happens before some operation $o' \in \textit{LMSet}_i(h,x)$ in $h$ $\}$. %We can see that for each $i$, $\textit{LMSet}_i(h,x)$ contains matched $\textit{put}$ and $\textit{rm}$ operations, and
We can see that $\textit{LMSet}_i(h,x) \cap \textit{LMSet}_j(h,x) = \emptyset$ for any $i \neq j$. Let $\textit{LMSet}(h,x) = \textit{LMSet}_1(h,x) \cup \textit{LMSet}_2(h,x) \cup \ldots$. %{\color {red} We say that a history $h$ of priority queue is matched before $x$, if (1) $h$ contains no $\textit{rmEmpty}$ operations, (2) $ h \sqsubseteq l_1 \cdot \textit{put}(x,\_) \cdot l_2 \cdot \textit{rm}(x) \cdot l_3$ for some $l_1,l_2,l_3$, and $l_1 \cdot l_2$ contains matched $\textit{put}$ and $\textit{rm}$ operations.}

The following is an example that shows that, given a data-differentiated history $h$, the set $\textit{LMSet}_i(h,x)$ can be non-empty for each $1 \leq i \leq 3$. In \figurename~\ref{fig:his nobound of LMSet} , $\textit{LMSet}_1(h,x) = \{ \textit{put}(a,1),\textit{rm}(a) \}$, $\textit{LMSet}_2(h,x) = \{ \textit{put}(b,1),\textit{rm}(b) \}$ and $\textit{LMSet}_3(h,x) = \{ \textit{put}(a,c),\textit{rm}(c) \}$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_HIS_NOBOUNDOF_LMSET.pdf}
%\vspace{-10pt}
  \caption{An example for non-emptiness of $\textit{LMSet}_1(h,x)$, $\textit{LMSet}_2(h,x)$ and $\textit{LMSet}_3(h,x)$}
  \label{fig:his nobound of LMSet}
\end{figure}

\begin{restatable}{lemma}{RmxDoesNotHappenBeforeLMSetForRpr1}
\label{lemma:Rmx does not happen before LMSet for Rpr1}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. %, such that $\textit{pri}_x$ is the maximal priority of $h$, and $\textit{last}(h) = R_{\textit{pr1}}$.
Assume that $\textit{rm}(y)$ does not happen before $\textit{put}(y,\_)$ for each item $y$ in $h$. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}(h,x)$.
\end{restatable}

We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$, in $\textit{LMSet}_2(h,x)$, $\ldots$. Note that, according to the definition of $\textit{LMSet}(h,x)$, $\textit{pri}_x$ is different to priority of all items in $\textit{LMSet}(h,x)$.

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_1(h,x)$ is an operation of item $d$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_1(h,x)$: $o \in \textit{LMSet}_1(h,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

@@

This completes the proof for $\textit{LMSet}_1(h,x)$.

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_{\textit{j+1}}(h,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$ is an operation of item $d_{\textit{j+1}}$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_{\textit{j+1}}$ does not equals $\textit{pri}_x$). We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ operation or a $\textit{rm}$ operation. $t_2$ and $t_3$ is used for the reason of $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$: $o \in \textit{LMSet}_{\textit{j+1}}(h,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

@@

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$. Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{LMSet}_j(h,x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(_{\textit{j-1}},\_), \textit{rm}(_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_j(h,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{LMSet}_{\textit{j+1}}(h,x)$.

@@

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{LMSet}_{\textit{j-1}}(h,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.

@@

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$. Here each $d_i$ is a item of some operation in $\textit{LMSet}_i(h,x)$ (according to the definition of $\textit{LMSet}_1(h,x)$, the priority of $d_i$ does not equals $\textit{pri}_x$ for each $i$). Let us prove that from $T_j$ we can obtain contradiction by induction:

@@

    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$. Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{LMSet}_1(h,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{LMSet}_2(h,x)$.
    \end{itemize}

@@

    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$. This implies that $\textit{put}(d_3,\_) \in \textit{LMSet}_2(h,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{LMSet}_3(h,x)$. Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.

@@

    By base case $1$, base case $2$ and the induction step, it is easy to see that $T_j$ contains contradiction.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{LMSet}_j(h,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{LMSet}_1(h,x) \cup \ldots \cup \textit{LMSet}_j(h,x)$.

>>>>

For each such case, we construct a witness automata. We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

@@

\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of PQ1Lar}}
\label{sec:appendix proof and definition in section co-regular of PQ1Lar}

We can see that $\textit{UVSet}_i(e,x) \cap \textit{UVSet}_j(e,x) = \emptyset$ for any $i \neq j$.

The following lemma states that $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.

\begin{restatable}{lemma}{UVSetHasMatchedPutandRm}
\label{lemma:UVSet has matched put and rm}
Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.
\end{restatable}
\begin {proof}

We prove this lemma by contradiction. Assume that there exists a value, such that $\textit{UVSet}(e,x)$ contains only its $\textit{put}$ and does not contain its $\textit{rm}$. Then we can see that there exists $d_1,\ldots,d_j$. Intuitively, $d_1,\ldots,d_j$ are elements in $\textit{UVSet}_1(e,x), \ldots, \textit{UVSet}_i(e,x)$, respectively. $\textit{UVSet}(e,x)$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$. And each $d_i$ is the reason of $d_{\textit{i+1}} \in \textit{UVSet}_{\textit{i+1}}(e,x)$. Formally, we require that

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For each $1 \leq i \leq j$, method events of $d_i$ belongs to $\textit{UVSet}_i(e,x)$.

\item[-] For each $i \neq j$, $\textit{put}(d_i,\_),\textit{rm}(d_i) \in \textit{UVSet}_i(e,x)$. $\textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x)$, and $e$ does not contain $\textit{rm}(d_j)$.

\item[-] An operation of $d_1$ happens before an operations of $x$. For each $1 < i \leq j$, an operation of $d_i$ happens an operation of $d_{\textit{i-1}}$.

According to the definition of $\textit{UVSet}(e,x)$, it is easy to see that such $d_1,\ldots,d_j$ exists. Let us prove the following fact:

Proof of $\textit{fact}_1$: We prove $\textit{fact}_1$ by contradiction. Assume that for some $i \neq j$, $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap. Since $\textit{put}(d_i,\_), \textit{rm}(d_i) \in \textit{UVSet}_i(h,x)$, we know that an operation $o_i$ of $d_i$ happens before operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. Moreover, since $\textit{put}(d_i,\_)$ and $\textit{rm}(d_i)$ overlap, it is not hard to see that the call action of $\textit{put}(d_i,\_)$ and the call action of $\textit{rm}(d_i)$ is before the call action of $o_{\textit{i-1}}$. Since method events of $d_{\textit{i+1}}$ is in $\textit{UVSet}_{\textit{i+1}}(e,x)$, we know that an operation $o'_{\textit{i+1}}$ of $d_{\textit{i+1}}$ happens before operation $o'_i$ of $d_i$. Then, it is not hard to see that $o'_{\textit{i+1}}$ also happens before $o_{\textit{i-1}}$, which contradicts that for each $k > \textit{ind+1}$, no operation of $d_k$ happens before operation of $d_{\textit{ind}}$.

We already know that an operation of $d_1$ happens before an operation of $x$. By $\textit{fact}_1$, we can ensure that $\textit{put}(d_1,\_)$ happens before an operation of $x$, and then $d_1 \rightarrow x$ in $G$. For each $1 < i \leq j$, we know that an operation $o_i$ of $d_i$ happens before an operation $o_{\textit{i-1}}$ of $d_{\textit{i-1}}$. By $\textit{fact}_1$, we can ensure that $o_i=\textit{put}(d_i,\_)$ and $o_{\textit{i-1}}=\textit{rm}(d_{\textit{i-1}})$, and then $d_i \rightarrow d_{\textit{i-1}}$ in $G$. Since $h$ contains $\textit{put}(d_j,\_)$ and does not contain $\textit{rm}(d_j)$, we know that $x \rightarrow d_j$ in $G$. Then $G$ has a cycle going through $x$, contradicts that $G$ has no cycle going through $x$. \qed
\end {proof}

The following lemma states that $\textit{UVSet}(e,x)$ does not happen before $\textit{rm}(x)$ when the left-right constraint has no cycle going through $x$.

\begin{restatable}{lemma}{RmxDoesNotHappenBeforeUVSetForPQ1Lar}
\label{lemma:Rmx does not happen before UVSet for PQ1Lar}

Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{>}$. Let $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ be method events of $e$ with maximal priority. Let $G$ be the graph representing the left-right constraint of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$. Assume that $G$ has no cycle going through $x$. Then, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}(e,x)$.
\end{restatable}

@@

We prove this lemma by induction, and prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$, in $\textit{UVSet}_2(e,x)$, $\ldots$. Note that, by Lemma \ref{lemma:UVSet has matched put and rm}, $\textit{UVSet}(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that for each $i$, $\textit{UVSet}_i(e,x)$ contains only matched $\textit{put}$ and $\textit{rm}$.

\noindent (1) Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_1(e,x)$ is an operation of item $d$. %(according to the definition of $\textit{UVSet}_1(h,x)$, the priority of $d$ does not equals $\textit{pri}_x$).

We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_1(e,x)$: $o \in \textit{UVSet}_1(e,x)$, since an operation (of kind $t_2$) of $d$ happens before an operation (of kind $t_3$) of $x$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

@@

This completes the proof for $\textit{UVSet}_1(e,x)$.

\noindent (2) Assume we already prove that for some $j \geq 1$, $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$. Let us prove that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_{\textit{j+1}}(e,x)$ by contradiction. Assume that $\textit{rm}(x) <_{hb} o$, where $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$ is an operation of item $d_{\textit{j+1}}$. We use a triple $(t_1,t_2,t_3)$ to represent related information. $t_1,t_2,t_3$ are chosen from $\{ \textit{put},\textit{rm} \}$. $t_1$ represents whether $o$ is a $\textit{put}$ method event or a $\textit{rm}$ method event. $t_2$ and $t_3$ is used for the reason of $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$: $o \in \textit{UVSet}_{\textit{j+1}}(e,x)$, since an operation (of kind $t_2$) of $d_{\textit{j+1}}$ happens before an operation (of kind $t_3$) of $d_j$, where $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$. Let us consider all the possible cases of $(t_1,t_2,t_3)$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $(\textit{put},\textit{put},\textit{put})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

\item[-] $(\textit{put},\textit{put},\textit{rm})$: Then $\textit{rm}(x) <_{hb} \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j,\_)$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j,\_) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

@@

\item[-] $(\textit{rm},\textit{put},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) )$. Let us consider the reason of $\textit{put}(d_j,\_), \textit{rm}(d_j) \in \textit{UVSet}_j(e,x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $( j > 1 ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) <_{hb} o'' )$, we can see that $\textit{put}(d_{\textit{j+1}},\_) <_{hb} o''$, and then operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_j(e,x)$, contradicts that operations of $d_{\textit{j+1}}$ is in $\textit{UVSet}_{\textit{j+1}}(e,x)$.

@@

    \item[-] If $( j > 1 ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, where $o''$ is an operation of item $d_{\textit{j-1}}$ and $\textit{put}(d_{\textit{j-1}},\_), \textit{rm}(d_{\textit{j-1}}) \in \textit{UVSet}_{\textit{j-1}}(e,x)$: Then since $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{rm}(d_j) <_{hb} o'' )$, we can see that $( \textit{put}(d_{\textit{j+1}},\_) <_{hb} o'' ) \vee ( \textit{rm}(d_j) <_{hb} \textit{put}(d_j,\_) )$, which is impossible.

@@

\item[-] $(\textit{rm},\textit{put},\textit{rm})$: Let $T_{\textit{ind}}$ be the set of sentences $\{ \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}), \textit{put}(d_{\textit{j+1}},\_) <_{hb} \textit{rm}(d_j),\ldots, \textit{put}(d_{\textit{ind+1}},\_) <_{hb} \textit{rm}(d_{\textit{ind}}) \}$. Here each $d_i$ is a item of some operation in $\textit{UVSet}_i(e,x)$. Let us prove that from $T_j$ we can obtain contradiction by induction:

@@

    \item[-] If $\textit{rm}(d_1)$ happens before $o$, and $o$ is an operation of $x$. Then since $\textit{put}(d_2,\_) <_{hb} \textit{rm}(d_1)$ and $\textit{rm}(d_1) <_{hb} o$, we can see that $\textit{put}(d_2,\_) <_{hb} o$, and then $\textit{put}(d_2,\_) \in \textit{UVSet}_1(e,x)$, contradicts that $\textit{put}(d_2,\_) \in \textit{UVSet}_2(e,x)$.
    \end{itemize}

@@

    Let us prove base case $2$: If $\textit{rm}(d_2) <_{hb} o$, and $o$ is an operation of $d_1$, then since $( \textit{put}(d_3,\_) <_{hb} \textit{rm}(d_2) ) \wedge ( \textit{rm}(d_2) <_{hb} o )$, we know that $\textit{put}(d_3,\_) <_{hb} o$. This implies that $\textit{put}(d_3,\_) \in \textit{UVSet}_2(e,x)$, contradicts that $\textit{rm}(d_3,\_) \in \textit{UVSet}_3(e,x)$. Therefore, it is only possible that $\textit{put}(d_2,\_)$ happens before an operation of $d_1$.

@@

    By base case $1$, base case $2$ and the induction step, it is easy to see that for each $i$, $T_i$ contains contradiction. Therefore, $T_j$, the case of $(\textit{rm},\textit{put},\textit{rm})$, contains contradiction.

\item[-] $(\textit{rm},\textit{rm},\textit{put})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{put}(d_j,\_) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{put}(d_j,\_) ) \wedge ( \textit{put}(d_j,\_) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.

\item[-] $(\textit{rm},\textit{rm},\textit{rm})$: Then $( \textit{rm}(x) <_{hb} \textit{rm}(d_{\textit{j+1}}) ) \wedge ( \textit{rm}(d_{\textit{j+1}}) <_{hb} \textit{rm}(d_j) )$. We can see that $( \textit{rm}(x) <_{hb} \textit{rm}(d_j) ) \wedge ( \textit{rm}(d_j) \in \textit{UVSet}_j(e,x) )$, which contradicts that $\textit{rm}(x)$ does not happen before any operation in $\textit{UVSet}_1(e,x) \cup \ldots \cup \textit{UVSet}_j(e,x)$.
\end{itemize}
